# This reader

```{r, include=FALSE}
knitr::opts_chunk$set(echo = T, class.source="Rchunk", class.output="Rout")
```

This reader has a number of conventions. Below, we indicate the most important ones.

## Exercises

###### example code along {.oranje}
"code along exercises" (one of the trainers will demo something while you copy their steps in your own github account) have an orange background. We will not actually write a lot of code by the way. Perhaps even none.




## Code blocks

This is Bash code, that can be run in a terminal or from a Bash script (`.sh`).
```
# Bash
ls -a
```

This is R code, and its output under the code.
```{r}
x <- c(1:10)
x
```

## Block quotes
We enclose code, file_names, sytem variables, commands or other code associated items with `block quotes`

## R Packages
When referring to R packages we use `{package_name}`

## Assignments
These can be recognized with a specific style, where the assinment is enclosed in a text-box. We created this mark-up with a CSS style sheet. See file `style.css` for details.
